Structured Versus Decorated Cospans

### 2020-11-26 cospans.tex
a first take on related work

In this section we present several examples.  The first was mentioned in the Introduction.

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

In this section we present several examples, drawing the relation between the formalisms of structured and decorated cospans. Before we exhibit these formalisms in action, let us briefly remark that this work is part of a more general endeavor to apply categorical ideas to the theory and analysis of systems in a broad sense. Various capproaches to different problems have been followed; thinking about systems and processes categorically dates back to early works like \cite{Lawvere,JNW,BF2,KSW} and references therein. As mentioned in the introduction, the current material constitutes part of the line of work initiated by Fong \cite{Fong,FongThesis} and evolved and applied in different settings through a number of publications (here cite papers/theses that surround decorated and structured cospans in a broad sense). A different line of work employs algebras of wiring diagrams and sheaves to capture compositional and time features of dynamical systems, initiated by Spivak and progressed also onto cyber-physical directions \cite{VSL,SSV,BFV}. (Another line of work could be props used for systems? John?) Another approach uses graphical languages and signal flow diagrams \cite{FRS,BPSZ} to understand systems of a more behavioral taste \cite{Willems}.

Having discussed some by no means exhaustive related work in the field, we are ready to set out to the examples of our framework.

@@

{\chris Shouldn't we give Shulman's result that the horizontal bicategory of a symmetric monoidal double category is symmetric monoidal?}

@@

\bibitem{BFV} G.\ Bakirtzis, C.\ H.\ Fleming and C.\ Vasilakopoulou, Categorical Semantics of Cyber-Physical Systems Theory. Available as 
\href{https://arxiv.org/abs/2010.08003}{arXiv:2010.08003}.

\bibitem{BPSZ} F.\ Bonchi, R.\ Piedeleu, P.\ Sobocinski and F.\ Zanasi, Contextual Equivalence for Signal Flow Graphs, \textsl{FoSSaCS} (2020). Available as \href{https://arxiv.org/abs/2002.08874}{arXiv:2002.08874}.

@@

\bibitem{BF2} M.\ Bunge and M.\ Fiore, Unique factorisation lifting functors and categories of linearly-controlled processes, \textsl{Math.\ Struct.\ Comput.\ Sci.\ } \textbf{10} (2000), 137--163.

@@

\bibitem{FRS} B.\ Fong, P.\ Rapisarda, P.\ Sobocinski, \textsl{A categorical approach to open and interconnected dynamical systems}, LICS, 2016. Available as \href{http://arxiv.org/abs/510.05076}{arXiv:1510.05076}.

@@

\bibitem{JNW} A.\ Joyal, M.\ Nielsen and G.\ Winskel, Bisimulation from Open Maps, \textsl{Information and Computation} \textbf{127} (1996) 164 -- 185.

\bibitem{KSW} P.\ Katis, N.\ Sabadini and R.\ F.\ C.\ Walters, On the algebra of systems with feedback and boundary, \textsl{Rendiconti del Circolo Matematico di Palermo Serie II} \textbf{63} (2000) 123 -- 156.

@@

\bibitem{Lawvere} F.\ W.\ Lawvere, State categories and response functors, unpublished manuscript (1986)

@@

\bibitem{SSV} P.\ Schultz, D.\ Spivak and C.\ Vasilakopoulou, Dynamical systems and sheaves, \textsl{Appl.\ Cat.\ Struct.\ }\textbf{28} (2020), 1--57. Available as \href{https://arxiv.org/abs/1609.08086}{arXiv:1609.08086}.

@@

\bibitem{VSL} D.\ Vagner, D.\ Spivak and E.\ Lerman, Algebras of open dynamical systems on the operad of wiring diagrams, \textsl{Theory Appl.\ Categ.\ } \textbf{30} (2015), 1793--1822. Available as \href{https://arxiv.org/abs/1408.1598}{arXiv:1408.1598}.

\bibitem{Willems} J.\ C.\ Willems, The behavioral approach to open and interconnected systems, \textsl{IEEE Control Systems Magazine} (2007)
